### 2023-09-03 dev/dev_template_report.Rmd
Template project rmarkdown report

---
title: Template report example
subtitle: subtitle
author:
  name: author
  email: email@example.com
  affiliation: "name <br> address"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# Libraries
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(streamFind)

# Variables
logo <- NA_character_

# Global options
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis")

# Working directory
path_wd <- getwd()

# Add logo file
path_extdata <- system.file(package = "streamFindData", dir = "extdata")
logo <- file.path(path_extdata, "logo_iuta.png") #207 x 216 px
```

```{r logo, echo=FALSE}
if (!is.na(logo)) {
  htmltools::img(
  src = knitr::image_uri(logo),
  alt = "logo",
  style =
    paste0("
      margin-left:inherit;
      position:absolute;
      top:60px;
      left:740px;
      width:", 100, "px;
      height:", 100*(216/207), "px;"
    )
  ) 
}
```

<br>

***

# Instructions

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
This section gives quick instructions to display text, code, tables and figures.

## Text

Text in R markdown has several inline codes for costumization.
The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.

## Code

Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
Code chunks should always be unically named.

```{r exampleCode}

#Example of code chunk

list(a = 1:6, b = 1:10)

```

<br>

## Tables

The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exampleTable). Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
Table caption is added with the argument `caption` of the `kable` function.

```{r exampleTable}

df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)

knitr::kable(df, caption = "Example of table caption.") %>%
kable_styling(
  font_size = 12,
  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
  fixed_thead = TRUE
) %>%
  scroll_box(width = "100%", height = "350px")

```

<br>

## Figures

Figures/plots can also by embedded using diverse packages.
The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.

### Example `ggplot2`

```{r examplePlot, fig.cap="Plot caption example"}

ggplot(pressure, aes(temperature, pressure)) +
geom_point() +
theme_bw()

```

### Example `plotly`

```{r examplePlotly, fig.cap="Plot caption example"}

plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")

```

<br> <br>

# Signature

The report was produced on `r Sys.time()`. The data and report are stored in ``r path_wd`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.

***

<br>

<style>
body, p {
  color: black;
  font-family: Arial;
  text-align: justify;
  font-size: 12pt;
}
h1{
  font-size: 20pt;
  font-style: bold;
}
h2{
  font-size:18pt;
  font-style: bold;
}
h3{
  font-size:14pt;
  font-style: bold;
}
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #239B56;
}
caption {
  color: black;
  font-size: 1.0em;
}
</style>