### 2024-01-05 dev/TiChri_integration_patRoon.Rmd
Improved as_fGroups

---
title: "TiChri: Matrix correction integration in patRoon"
subtitle: "Concept"
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
  affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Libraries
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)
library(patRoon)
# Global options
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis", cache = TRUE)
options(width = 300)
```

```{r logo, echo=FALSE}
path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
logo <- file.path(path_extdata, "logo_StreamFind.png")
htmltools::img(src = knitr::image_uri(logo), style = "margin-left:inherit;position:absolute;top:80px;left:740px;width:200px;")
```

<br>

***

```{r resources, include=FALSE}
files <- list.files("D:/NTS/Project_220614_AirLiquid_220509_SPE/msfiles", full.names = TRUE)
selected_files <- files[grepl("pos_Blank|pos_M220509014|pos_M220509017|pos_M220509018", files)]

all_db <- StreamFindData::get_ms_tof_spiked_chemicals_with_ms2()
db <- all_db[!grepl("IS", all_db$tag, fixed = TRUE), ]
cols <- c("name", "formula", "mass", "SMILES", "rt", "polarity", "fragments")
# cols <- c("name", "formula", "mass", "SMILES", "rt")
db <- db[, cols, with = FALSE]
# data.table::setnames(db, "mass", "neutralMass")

# db$neutralMass <- db$neutralMass + 1.0073
# data.table::setnames(db, "neutralMass", "mz")

dbis <- all_db[grepl("IS", all_db$tag), ]
cols <- c("name", "formula", "mass", "rt")
dbis <- dbis[, cols, with = FALSE]
# files <- list.files("E:/Pub_Aopti/ms_files_mzML", "M200422", full.names = TRUE)
```

# AnalysisInfo

```{r anaInfo}
# The 3 first files are control and were excluded
anaInfo <- data.frame(
  "path" = dirname(selected_files),
  "analysis" = tools::file_path_sans_ext(basename(selected_files)),
  "group" = c(
    rep("blank", 2),
    rep("effluent_OZ", 2),
    rep("effluent_OSW", 2),
    rep("influent", 2)
  ),
  "blank" = rep("blank", 8)
)

kable(anaInfo, caption = "AnalysisInfo") %>%
  kable_styling(
    font_size = 12,
    bootstrap_options = c("striped", "hover", "condensed"),
    fixed_thead = TRUE
  )
```

<br>

# Total ion Chromatograms (TICs)

```{r tics, message=FALSE, fig.height=9, fig.cap="Total ion chromatogram (TIC) of all analyses by replicate group name."}
ms <- MassSpecData$new(selected_files)
ms$add_replicate_names(anaInfo$group)
ms$add_blank_names(anaInfo$blank)
ms$plot_tic(levels = 1, colorBy = "replicates")
```

<br>

# Data processing

Processing settings described in \@ref(processing-settings).

```{r data-process, message=FALSE, results='markup'}
ms$find_features(Settings_find_features_openms())
ms$group_features(Settings_group_features_openms())
ms$find_internal_standards(Settings_find_internal_standards_StreamFind(database = dbis, ppm = 8, sec = 10))
ms
```

```{r internal-standards, fig.height=9, fig.cap="Evaluaton of spiked internal standards after feature finding and grouping."}
#ms$plot_internal_standards_qc()
```

```{r to-patroon, message=FALSE, results='markup'}
fGroups <- ms$as_patRoon_featureGroups()
fGroups
```

<br>

# Implementation in patRoon

## Get TICs

```{r get-tics-func, warning=FALSE, message=FALSE}
getTICs <- function(anaInfo) {
  filePaths <- patRoon:::getMzMLOrMzXMLAnalysisPath(anaInfo$analysis, anaInfo$path, mustExist = TRUE)
  # can be applied in parallel?
  tics <- lapply(filePaths, function(fpath) {
    msf <- mzR::openMSfile(fpath)
    hd <- mzR::header(msf)
    mzR::close(msf)
    hd <- hd[hd$msLevel == 1, ]
    data.table::data.table("rt" = hd$retentionTime, "int" = hd$totIonCurrent)
  })
  names(tics) <- anaInfo$analysis
  tics
}
```

```{r get-tics, warning=FALSE, message=FALSE, results='markup'}
tics <- getTICs(anaInfo)

length(tics)

is(tics)

head(tics[[1]])
```

## Calculate TIC matrix profiles (MPs)

The function uses a retention window (`rtWindow`), in seconds, to estimate the blank level in the proximity of the analysis spectrum. For now, simple average is used to aggregate the TIC MPs calculated for each blank replicate group associated with a given analysis and then to summarize the analysis MPs into one matrix profile (MP) vector. The MP is calculated according to eq. 6 (see below) of 10.1021/acs.analchem.1c00357.

$TIC_{MP} = (\frac{TIC^{matrix}}{TIC^{blank}}) * -1$

```{r calc-mps-func, warning=FALSE, message=FALSE}
calculateTicMps <- function(anaInfo, tics, rtWindow = 10) {

  .trimVector <- function(v, a, b) {
    rowSums(mapply(function(a, b) v >= a & v <= b, a = a, b = b)) > 0
  }

  blkGroups <- anaInfo$blank
  names(blkGroups) <- anaInfo$analysis

  blkAnalyses <- anaInfo$analysis
  names(blkAnalyses) <- anaInfo$blank
  blkAnalyses <- blkAnalyses[anaInfo$group %in% anaInfo$blank]

  # can be applied in parallel?
  anaMpList <- lapply(anaInfo$analysis, function(x, blkGroups, blkAnalyses, tics, rtWindow) {

    blkG <- blkGroups[x]
    blkA <- blkAnalyses[names(blkAnalyses) %in% blkG]

    ticX <- tics[[x]]

    mpListX <- lapply(tics[blkA], function(z, ticX, rtWindow) {

      mp <- vapply(seq_len(nrow(ticX)), function(j, z) {
        rt <- ticX$rt[j]
        rtr <- c(rt - rtWindow, rt + rtWindow)
        intB <- mean(z$int[.trimVector(z$rt, rtr[1], rtr[2])])
        # eq. 6 from 10.1021/acs.analchem.1c00357
        (ticX$int[j] / intB) * -1
      }, z = z, 0)

      mp

    }, ticX = ticX, rtWindow = rtWindow)

    ticX$mp <- Reduce(`+`, mpListX) / length(mpListX)

    ticX

  }, blkGroups = blkGroups, blkAnalyses = blkAnalyses, tics = tics, rtWindow = rtWindow)

  names(anaMpList) <- anaInfo$analysis

  anaMpList
}
```

```{r calc-mps, warning=FALSE, message=FALSE, fig.height=9, fig.cap="MPs of each analysis"}
ticMps <- calculateTicMps(anaInfo, tics, 10)

fig <- plot_ly()
cl <- StreamFind:::.get_colors(anaInfo$group)
sL <- !duplicated(cl)
for(i in seq_len(nrow(anaInfo))) {
  fig <- fig %>% add_trace(
    x = ticMps[[i]]$rt,
    y = ticMps[[i]]$mp,
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = sL[i],
    type = "scatter",
    mode = 'lines',
    line = list(color = cl[i])
  )
}

# fig <- fig %>% add_trace(
#     x = iseval$rt,
#     y = iseval$rec,
#     name = names(cl)[i],
#     legendgroup = names(cl)[i],
#     showlegend = sL[i],
#     type = "scatter",
#     mode = 'markers',
#     marker = list(color = cl[i])
#   )

xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Retention time / seconds",
  titlefont = list(size = 12, color = "black")
)

yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Suppression factor",
  titlefont = list(size = 12, color = "black")
)

fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)

fig
```

```{r}

iseval <- ms$get_internal_standards(average = T)

plot(iseval$rt, iseval$rec)

```

s
```{r}
purine <- 121.05090000 
HP921 <- 922.00980000

# ms$plot_eic(analyses = c(2, 5, 14), mass = dbis$mass[4], ppm = 20, colorBy = "targets+analyses")

ms$plot_eic(analyses = c(2, 5, 14), mass = c(purine, HP921), ppm = 20, colorBy = "replicates+targets")

# Check a project with SPE applied
```

```{r}

```

<br>

# Processing settings

```{r pro-settings, warning=FALSE, message=FALSE, results='markup'}
Settings_find_features_openms()
Settings_group_features_openms()
```