### 2024-05-06 dev/dev_paracetamol.Rmd
Added workflow for paracetamol in dev

---
title: 'Fusion of Mass Spectrometry and Raman Spectroscopy Data'
subtitle: |
  | Comparison of paracetamol deviations
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Fusion of Mass Spectrometry and Raman Spectroscopy Data}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", fig.width = 9, results = "asis", comment = "")

main_dir <- "C:/Users/apoli/Documents/iSoft/Paracetamol"

ms_dir <- paste0(main_dir, "/ms/mzml")

ms_files <- list.files(ms_dir, pattern = ".mzML", full.names = TRUE)

raman_dir <- paste0(main_dir, "/raman")

raman_files <- list.files(raman_dir, pattern = ".asc", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

# Introduction

# Pre-processing

## MS

```{r}
ms <- MassSpecEngine$new(files = ms_files, headers = list(name = "BSA Quality Evaluation Project"))

ms$add_replicate_names(
  c(rep("fiz_1", 3),
    rep("fiz_2", 3),
    rep("fiz_3", 3),
    rep("set_32_1", 3),
    rep("set_32_2", 3),
    rep("set_32_3", 3),
    rep("set_33_1", 3),
    rep("set_33_2", 3),
    rep("set_33_3", 3),
    rep("fiz_caff_1", 3),
    rep("fiz_caff_2", 3),
    rep("fiz_caff_3", 3),
    rep("fiz_ass_1", 3),
    rep("fiz_ass_2", 3),
    rep("fiz_ass_3", 3),
    rep("fiz2_1", 3),
    rep("fiz2_2", 3),
    rep("fiz2_3", 3),
    rep("fiz_acid_1", 3),
    rep("fiz_acid_2", 3),
    rep("fiz_acid_3", 3),
    rep("ass_1", 3),
    rep("ass_2", 3),
    rep("ass_3", 3),
    rep("ass2_1", 3),
    rep("ass2_2", 3),
    rep("ass2_3", 3),
    rep("caff_1", 3),
    rep("caff_2", 3),
    rep("caff_3", 3),
    rep("par_1", 3),
    rep("par_2", 3),
    rep("par_3", 3),
    rep("blank", 23)
  )
)

# View(ms$get_overview())
```

```{r, fig.cap="Total ion chromatograms for each analysis replicate."}
# ms$plot_spectra_tic(colorBy = "replicates")

# only blanks
# ms$plot_spectra_tic(analyses = 100:122, colorBy = "analyses")
```

```{r}
ms <- ms$subset_analyses(analyses = c(1:9, 28:72, 82:99, 118:121))
ms$add_blank_names(rep("blank", 76))
ms$load_spectra()
```

```{r}
ms_workflow_settings <- list(
  Settings_bin_spectra_StreamFind(bins = list("rt" = 5, "mz" = 2), refBinAnalysis = 1),
  Settings_normalize_spectra_minmax(),
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(negativeToZero = TRUE),
  Settings_normalize_spectra_blockweight()
)

ms$add_settings(ms_workflow_settings)

ms$print_workflow()
```

```{r}
ms$run_workflow()
```

```{r ,fig.cap="Deconvoluted and pre-processed spectra for each analysis."}
ms$plot_spectra()
```

## Raman

```{r}
raman <- RamanEngine$new(raman_files)

raman$add_replicate_names(
  c(
    rep("ass_1", 3),
    rep("ass_2", 3),
    rep("ass_3", 3),
    rep("caff_1", 3),
    rep("caff_2", 3),
    rep("caff_3", 3),
    rep("blank", 9),
    rep("fiz_1", 3),
    rep("fiz_2", 3),
    rep("fiz_3", 3),
    rep("fiz_ass_1", 3),
    rep("fiz_ass_2", 3),
    rep("fiz_ass_3", 3),
    rep("fiz_ass_b_2", 9),
    rep("fiz_caff_1", 3),
    rep("fiz_caff_2", 3),
    rep("fiz_caff_3", 3),
    rep("fiz_b", 9),
    rep("fiz2_1", 3),
    rep("fiz2_2", 3),
    rep("fiz2_3", 3),
    rep("fiz_d", 9),
    rep("background", 4),
    rep("fiz_acid_1", 3),
    rep("fiz_acid_2", 3),
    rep("fiz_acid_3", 3),
    rep("par_1", 3),
    rep("par_odd", 3),
    rep("par_3", 3),
    rep("par_2", 3),
    rep("cali_150_mgml", 9),
    rep("cali_010_mgml", 9),
    rep("cali_025_mgml", 9),
    rep("cali_050_mgml", 9),
    rep("cali_075_mgml", 9),
    rep("background", 4)
  )
)

# View(raman$get_overview())
```

```{r ,fig.cap="Raw averaged Raman spectra for each BSA product between minute 5.5 and 8."}
# raman$plot_spectra(colorBy = "replicates")
```

```{r}
raman <- raman$subset_analyses(analyses = c(1:45, 55:63, 73:81, 95:106, 110:115))
raman$add_blank_names(rep("blank", 81))
```

```{r}
raman_workflow_settings <- list(
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(),
  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 350))),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2500))),
  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30),
  Settings_normalize_spectra_minmax(),
  Settings_normalize_spectra_blockweight()
)

raman$add_settings(raman_workflow_settings)

raman$print_workflow()
```

```{r}
raman$run_workflow()
```

```{r ,fig.cap="Processed Raman spectra for each BSA product."}
raman$plot_spectra()
```

# Fusion

```{r}
rSpec <- raman$get_spectra()
rSpec <- rSpec[order(rSpec$replicate), ]
mrSpec <- matrix(rSpec$intensity, nrow = length(unique(rSpec$replicate)), ncol = length(unique(rSpec$shift)), byrow = TRUE, dimnames = list(as.character(unique(rSpec$replicate)), as.character(unique(rSpec$shift))))

mSpec <- ms$get_spectra()
mSpec <- mSpec[order(mSpec$replicate), ]
mmSpec <- matrix(mSpec$intensity, nrow = length(unique(mSpec$replicate)), ncol = length(unique(mSpec$bins)), byrow = TRUE, dimnames = list(unique(mSpec$replicate), unique(mSpec$bins)))

fusedMat <- cbind(mmSpec, mrSpec)

attr(fusedMat, "xaxis.name") = "Keys"
attr(fusedMat, "xaxis.values") = seq_len(ncol(fusedMat))
```

```{r}
cl <- StreamFind:::.get_colors(rownames(fusedMat))
fig <- plot_ly()
xVal <- seq_len(length(fusedMat[1, ]))
for (i in seq_len(nrow(fusedMat))) {
  fig <- fig %>% add_trace(
    x = xVal,
    y = fusedMat[i, ],
    type = "scatter", mode = "lines",
    line = list(width = 0.5, color = unname(cl[i])),
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = TRUE
  )
}
xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Bins",
  titlefont = list(size = 12, color = "black")
)
yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Intensity",
  titlefont = list(size = 12, color = "black")
)
fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
fig
```

# Statistic analysis (via mdatools)

### PCA

```{r, include=FALSE}
library(mdatools)
data(people)

m = pca(people, 7, scale = TRUE, info = "People PCA model")
m = selectCompNum(m, 5)

print(m$res$cal)

par(mfrow = c(1, 2))
mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
mdaplot(m$loadings, type = "p", show.labels = TRUE, show.lines = c(0, 0))
```

```{r}
fusedMat2 <- prep.autoscale(fusedMat, center = FALSE, scale = TRUE)
cl <- StreamFind:::.get_colors(rownames(fusedMat2))
fig <- plot_ly()
xVal <- seq_len(length(fusedMat2[1, ]))
for (i in seq_len(nrow(fusedMat2))) {
  fig <- fig %>% add_trace(
    x = xVal,
    y = fusedMat2[i, ],
    type = "scatter", mode = "lines",
    line = list(width = 0.5, color = unname(cl[i])),
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = TRUE
  )
}
xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Bins",
  titlefont = list(size = 12, color = "black")
)
yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Intensity",
  titlefont = list(size = 12, color = "black")
)
fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
fig
```

```{r, include=FALSE}

m = pca(fusedMat, 3, scale = FALSE, info = "People PCA model")
plot(m, show.labels = TRUE)
mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))

par(mfrow = c(2, 2))
plotVariance(m, type = "h", show.labels = TRUE)
plotQDoF(m)
plotT2DoF(m)
plotDistDoF(m)

par(mfrow = c(1, 2))
plotQDoF(m, main = "DoF plot (classic)")

m = setDistanceLimits(m, lim.type = "ddrobust")
plotQDoF(m, main = "DoF plot (robust)")

plotExtreme(m, comp = 1:2, main = "Extreme plot (cal)")

plotLoadings(m, 1:2, type = 'l')

# par(mfrow = c(2, 2))
# mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
# mdaplot(m$loadings, type = "p", show.labels = TRUE, show.lines = c(0, 0))
# plotResiduals(m$res$cal, show.labels = TRUE)
# plotVariance(m$res$cal, type = "h", show.labels = TRUE, labels = "values")
```

```{r ,echo=FALSE,fig.width=10,fig.height=4}

library(plotly)

loadings_plot <- plot_ly()

loadings_plot <- loadings_plot %>% add_trace(
  x = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 1"],
  y = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mrSpec))],
  textfont = list(size = 14),
  legendgroup = "Raman",
  name = "Raman",
  marker = list(size = 10)
)

loadings_plot <- loadings_plot %>% add_trace(
  x = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 1"],
  y = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mmSpec))],
  textfont = list(size = 14),
  legendgroup = "MS",
  name = "MS",
  marker = list(size = 10, color = "red")
)

loadings_plot %>% layout(
    title = "PCA Loadings Plot",
    xaxis = list(title = "Principal Component 1"),
    yaxis = list(title = "Principal Component 2")
  )

```

### PLS

```{r}

library(mdatools)
data(people)

idx = seq(4, 32, 4)
Xc = people[-idx, -4]
yc = people[-idx, 4, drop = FALSE]
Xt = people[idx, -4]
yt = people[idx, 4, drop = FALSE]

m = pls(Xc, yc, 7, scale = TRUE, cv = 1, info = "Shoesize prediction model")

par(mfrow = c(2, 2))
plotRegcoeffs(m)
plotRegcoeffs(m, ncomp = 2)
plot(m$coeffs, ncomp = 3, type = "b", show.labels = TRUE)
plot(m$coeffs, ncomp = 2)

par(mfrow = c(1, 2))
plotPredictions(m)
plotPredictions(m, ncomp = 1)

plotVIPScores(m, show.labels = TRUE)

res = predict(m, Xt, yt)
print(res)

data(simdata)
Xc = simdata$spectra.c
yc = simdata$conc.c[, 1]
Xt = simdata$spectra.t
yt = simdata$conc.t[, 1]

m = pls(Xc, yc, 10, x.test = Xt, y.test = yt, ncomp.selcrit = "min")
summary(m)

plot(m)

```

### PLS-DA

```{r}
data(iris)

cal.ind = c(1:25, 51:75, 101:125)
val.ind = c(26:50, 76:100, 126:150)

Xc = iris[cal.ind, 1:4]
Xv = iris[val.ind, 1:4]

cc.all = iris[cal.ind, 5]
cv.all = iris[val.ind, 5]

cc.vir = cc.all == "virginica"
cv.vir = cv.all == "virginica"
show(cc.vir)

m.all = plsda(Xc, cc.all, 3, cv = 1)
m.vir = plsda(Xc, cc.vir, 3, cv = 1, classname = "virginica")

summary(m.all)

par(mfrow = c(1, 2))
plotPredictions(m.all)
plotPredictions(m.vir)

res = predict(m.all, Xv, cv.all)
summary(res)

par(mfrow = c(1, 1))
plotPredictions(res, show.labels = TRUE)

```

### SIMCA
```{r}
idx = seq(1, nrow(iris), by = 2)

# split the values
Xc = iris[idx, 1:4]
cc = iris[idx, 5]

Xt = iris[-idx, 1:4]
ct = iris[-idx, 5]
```

### MCR

```{r}

data(carbs)
m = mcrpure(carbs$D, ncomp = 3, offset = 0.10)
summary(m)

par(mfcol = c(2, 2))
plotContributions(m)
plotSpectra(m)
plotSpectra(m, comp = 2)
plotContributions(m, comp = 2)

S = prep.norm(mda.t(carbs$S), "length")
S.hat = prep.norm(mda.t(m$resspec), "length")

c = predict(m, mda.t(carbs$S))
c = c / apply(c, 2, sum)
show(c)
```

```{r}
m = mcrpure(fusedMat, ncomp = 2, offset = 0.10)
summary(m)

par(mfcol = c(2, 2))
plotContributions(m)
plotSpectra(m)
plotSpectra(m, comp = 1)
plotContributions(m, comp = 1)

par(mfcol = c(1, 2))
plotVariance(m)
plotCumVariance(m)

plotPuritySpectra(m, ylim = c(0, 1600))

```

## Other Methods

### PCA in base R

```{r , include=FALSE, fig.width=10,fig.height=5}
data_pca <- prcomp(fusedMat, center = TRUE, scale. = FALSE)
pcs <- data_pca$x
pc_data <- data.frame(PC1 = pcs[, 1], PC2 = pcs[, 2], Species = rownames(pcs))
pca_plot <- plot_ly(
  data = pc_data,
  x = ~PC1,
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",
  text = ~Species,
  textfont = list(size = 14),
  marker = list(size = 12),
  textposition = "top center"
) %>%
  layout(
    title = "PCA Plot with Annotations",
    xaxis = list(title = "Principal Component 1"),
    yaxis = list(title = "Principal Component 2")
  )
pca_plot
```

```{r ,echo=FALSE,fig.width=10,fig.height=4}
loadings <- as.data.frame(data_pca$rotation[, 1:2])

loadings_plot <- plot_ly()

loadings_plot <- loadings_plot %>% add_trace(
  x = loadings[rownames(loadings) %in% colnames(mrSpec), "PC1"],
  y = loadings[rownames(loadings) %in% colnames(mrSpec), "PC2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(loadings)[(rownames(loadings) %in% colnames(mrSpec))],
  textfont = list(size = 14),
  legendgroup = "Raman",
  name = "Raman",
  marker = list(size = 10)
)

loadings_plot <- loadings_plot %>% add_trace(
  x = loadings[rownames(loadings) %in% colnames(mmSpec), "PC1"],
  y = loadings[rownames(loadings) %in% colnames(mmSpec), "PC2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(loadings)[(rownames(loadings) %in% colnames(mmSpec))],
  textfont = list(size = 14),
  legendgroup = "MS",
  name = "MS",
  marker = list(size = 10, color = "red")
)

loadings_plot %>% layout(
    title = "PCA Loadings Plot",
    xaxis = list(title = "Principal Component 1"),
    yaxis = list(title = "Principal Component 2")
  )

```

### Correlation Analysis

```{r}
# Calculate correlation matrix
correlation_matrix <- cor(fusedMat)

# Print correlation matrix
print(correlation_matrix)
```

### Distance Metrics

```{r}
# Calculate Euclidean distance matrix
distance_matrix <- dist(fusedMat)

# Print distance matrix
print(distance_matrix)
```

### Hierarchical Clustering

```{r}
# Perform hierarchical clustering
hc_result <- hclust(dist(fusedMat))

# Plot dendrogram
plot(hc_result)
```

### Multidimensional Scaling (MDS)

```{r}
# Calculate Euclidean distance matrix
distance_matrix <- dist(fusedMat)

# Perform MDS
mds_result <- cmdscale(distance_matrix)

# Plot MDS results
plot(mds_result[,1], mds_result[,2], pch = 19, col = "blue", xlab = "MDS1", ylab = "MDS2")

```

### Kernel Methods

```{r}
# Example code for kernel PCA (assuming Gaussian kernel)
library(kernlab)
kpca_result <- kpca(fusedMat, kernel = "rbfdot")

# Plot kernel PCA results
plot(kpca_result@pcv[,1], kpca_result@pcv[,2], pch = 19, col = "blue", xlab = "KPCA1", ylab = "KPCA2")
```

### Feature Selection and Importance

```{r}
# Example code for feature selection (using random forest as an example)
library(randomForest)
rf_model <- randomForest(fusedMat, importance = TRUE)
print(importance(rf_model))

var_importance <- importance(rf_model)
var_names <- rownames(var_importance)
var_scores <- var_importance[,"MeanDecreaseAccuracy"]  # Adjust column name as needed

# Sort variables by importance score
sorted_indices <- order(var_scores, decreasing = TRUE)
sorted_var_names <- var_names[sorted_indices]
sorted_var_scores <- var_scores[sorted_indices]

# Create bar plot
barplot(sorted_var_scores, names.arg = sorted_var_names, las = 2, cex.names = 0.8,
        main = "Variable Importance", xlab = "Variables", ylab = "Importance Score")
```

###  Cosine similarity and k-means clustering

```{r}
vectors <- fusedMat

similarity_matrix <- tcrossprod(vectors) / (sqrt(rowSums(vectors^2)) %*% t(sqrt(rowSums(vectors^2))))

# Perform k-means clustering
k <- 3  # Number of clusters
clusters <- kmeans(similarity_matrix, centers = k)

# Print cluster assignments
print(clusters$cluster)
```

### Similarity Measures

```{r}
vectors <- fusedMat

# Calculate cosine similarity between pairs of vectors
cosine_similarity <- function(x, y) {
  return(sum(x * y) / (sqrt(sum(x^2)) * sqrt(sum(y^2))))
}

similarity_matrix <- matrix(NA, nrow = nrow(vectors), ncol = nrow(vectors))
for (i in 1:nrow(vectors)) {
  for (j in 1:nrow(vectors)) {
    similarity_matrix[i, j] <- cosine_similarity(vectors[i,], vectors[j,])
  }
}

# Print similarity matrix
print(similarity_matrix)
```

### k-nearest neighbors (KNN) classification

```{r}
training_data <- fusedMat[c(1, 4), ]  # Replace this with your actual training data
test_data <- fusedMat[2:3, ]  # Replace this with your actual test data
labels <- rownames(fusedMat[c(1, 4), ])  # Replace this with your actual labels for training data

# Train KNN model
library(class)
knn_model <- knn(train = training_data, test = test_data, cl = labels, k = 3)

# Print predicted labels for test data
print(knn_model)
```

## Partial Least Squares (PLS)

Not applicable!

```{r}
# Example data
X <- fusedMat  # Replace this with your actual predictor variables
Y <- seq_len(nrow(fusedMat))  # Replace this with your actual response variable

# Perform PLS regression
library(pls)
pls_model <- plsr(Y ~ X, ncomp = 2)  # You can adjust the number of components (ncomp) as needed

# Predict using the PLS model
predicted_values <- predict(pls_model, newdata = X)

# Print predicted values
print(predicted_values)
```

## Support Vector Regression (SVR)

Not applicable!

```{r}
# Example data
X <- matrix(rnorm(100), nrow = 10)  # Replace this with your actual predictor variables
Y <- rnorm(10)  # Replace this with your actual response variable

# Train SVR model
library(e1071)
svr_model <- svm(Y ~ ., data = as.data.frame(X), kernel = "radial", epsilon = 0.1, cost = 1)  # Adjust parameters as needed

# Predict using the SVR model
predicted_values <- predict(svr_model, newdata = as.data.frame(X))

# Print predicted values
print(predicted_values)

```

## MCR-ALS (Multivariate Curve Resolution - Alternating Least Squares)

```{r}

# Perform MCR-ALS analysis
mcr_result <- mcr(fusedMat, nfac = 2)  # You can adjust the number of components (nfac) as needed

# Print MCR-ALS results
print(mcr_result)

```