### 2024-06-12 dev/demo_paracetamol.Rmd
Update articles

---
title: 'Detection of quality deviations in paracetamol products'
subtitle: |
  | Fusion of mass spectrometry and raman spectroscopy data
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Fusion of Mass Spectrometry and Raman Spectroscopy Data}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", fig.width = 9, results = "asis", comment = "")

main_dir <- "E:/iSoft/Paracetamol" # StreamFind Workstation

#main_dir <- "C:/Users/apoli/Documents/iSoft/Paracetamol" # Ricardo laptop

ms_dir <- paste0(main_dir, "/ms/mzml")

ms_dir_extra <- paste0(main_dir, "/ms/extra/mzml")

ms_files <- list.files(ms_dir, pattern = ".mzML", full.names = TRUE)

ms_files_extra <- list.files(ms_dir_extra, pattern = ".mzML", full.names = TRUE)

raman_dir <- paste0(main_dir, "/raman")

raman_dir_extra <- paste0(main_dir, "/raman/extra")

raman_files <- list.files(raman_dir, pattern = ".asc", full.names = TRUE)

raman_files_extra <- list.files(raman_dir_extra, pattern = ".asc", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

# Introduction

# Pre-processing

## MS

```{r ms-setup, include=FALSE}
ms <- MassSpecEngine$new(files = c(ms_files, ms_files_extra[c(5:40, 49:60)]), headers = list(name = "BSA Quality Evaluation Project"))

ms$add_replicate_names(
  c(rep("fiz_a1", 3),
    rep("fiz_a2", 3),
    rep("fiz_a3", 3),
    rep("fiz_b1", 3),
    rep("fiz_b2", 3),
    rep("fiz_b3", 3),
    rep("fiz_c1", 3),
    rep("fiz_c2", 3),
    rep("fiz_c3", 3),
    rep("fiz_caff1", 3),
    rep("fiz_caff2", 3),
    rep("fiz_caff3", 3),
    rep("fiz_ass1", 3),
    rep("fiz_ass2", 3),
    rep("fiz_ass3", 3),
    rep("fiz_d1", 3),
    rep("fiz_d2", 3),
    rep("fiz_d3", 3),
    rep("fiz_acid1", 3),
    rep("fiz_acid2", 3),
    rep("fiz_acid3", 3),
    rep("ass_n_1", 3),
    rep("ass_n_2", 3),
    rep("ass_n_3", 3),
    rep("ass_1", 3),
    rep("ass_2", 3),
    rep("ass_3", 3),
    rep("caff_1", 3),
    rep("caff_2", 3),
    rep("caff_3", 3),
    rep("par_1", 3),
    rep("par_2", 3),
    rep("par_3", 3),
    rep("blank", 23),
    rep("fiz_e01", 3),
    rep("fiz_e02", 3),
    rep("fiz_e03", 3),
    rep("fiz_e04", 3),
    rep("fiz_e05", 3),
    rep("fiz_e06", 3),
    rep("fiz_e07", 3),
    rep("fiz_e08", 3),
    rep("fiz_e09", 3),
    rep("fiz_e10", 3),
    rep("fiz_e11", 3),
    rep("fiz_e12", 3),
    rep("blank_e", 12)
  )
)

ms <- ms$subset_analyses(analyses = c(1:63, 73:99, 118:121, 123:130, 134:170))

ms$add_blank_names(c(rep("blank", 94), rep("blank_e", 139 - 94)))

ms$load_spectra()
```

```{r ms-overview, echo=FALSE}
ov <- ms$get_overview()[, 1:3]
setorder(ov, replicate)
ov$description <- c(
  rep("Aspirine standard", 9),
  rep("Reference background", 4),
  rep("Reference background", 9),
  rep("Caffein", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with acetic acid", 9),
  rep("Paracetamol with aspirine", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with caffein", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 36),
  rep("Paracetamol standard", 9)
)

kable(ov, caption = "Overview of MS analyses and respective analysis replicate groups and blanks.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r plot-ms-tic, fig.cap="Total ion chromatogram (TIC) for each analysis replicate."}
ms$plot_spectra_tic(analyses = ov$analysis[ov$replicate %in% "blank_e"], colorBy = "analyses", interactive = TRUE)
```

```{r ms-workflow}
ms_workflow_settings <- list(
  Settings_bin_spectra_StreamFind(bins = list("rt" = 5, "mz" = 2), refBinAnalysis = 1),
  Settings_normalize_spectra_minmax(),
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(negativeToZero = TRUE),
  Settings_normalize_spectra_blockweight()
)

ms$add_settings(ms_workflow_settings)

ms$print_workflow()
```

```{r ms-run-workflow}
ms$run_workflow()
```

```{r plot-ms-spectra, fig.cap="Pre-processed spectra for each analysis. Retention time dimension in he created bins is collapsed for each mass-to-charge ratio"}
ms$plot_spectra()
```

```{r ms-spectra}
ms$get_spectra()[1:10, ]
```

```{r}
ms$get_spectra_matrix()
```

## Raman

```{r raman-setup, include=FALSE}
raman_all <- RamanEngine$new(c(raman_files, raman_files_extra))

raman_all$add_replicate_names(
  c(
    rep("ass_1", 3),
    rep("ass_2", 3),
    rep("ass_3", 3),
    rep("caff_1", 3),
    rep("caff_2", 3),
    rep("caff_3", 3),
    rep("blank", 9),
    rep("fiz_c1", 3),
    rep("fiz_c2", 3),
    rep("fiz_c3", 3),
    rep("fiz_ass_n_1", 3),
    rep("fiz_ass_n_2", 3),
    rep("fiz_ass_n_3", 3),
    rep("fiz_ass1", 3),
    rep("fiz_ass2", 3),
    rep("fiz_ass3", 3),
    rep("fiz_caff1", 3),
    rep("fiz_caff2", 3),
    rep("fiz_caff3", 3),
    rep("fiz_a1", 3),
    rep("fiz_a2", 3),
    rep("fiz_a3", 3),
    rep("fiz_d1", 3),
    rep("fiz_d2", 3),
    rep("fiz_d3", 3),
    rep("fiz_b1", 3),
    rep("fiz_b2", 3),
    rep("fiz_b3", 3),
    rep("background", 4),
    rep("blank_e", 9),
    rep("fiz_acid1", 3),
    rep("fiz_acid2", 3),
    rep("fiz_acid3", 3),
    rep("fiz_e01", 3),
    rep("fiz_e02", 3),
    rep("fiz_e03", 3),
    rep("fiz_e04", 3),
    rep("fiz_e05", 3),
    rep("fiz_e06", 3),
    rep("fiz_e07", 3),
    rep("fiz_e08", 3),
    rep("fiz_e09", 3),
    rep("fiz_e10", 3),
    rep("fiz_e11", 3),
    rep("fiz_e12", 3),
    rep("par_1", 3),
    rep("par_odd", 3),
    rep("par_3", 3),
    rep("par_2", 3),
    rep("cali_150_mgml", 9),
    rep("cali_010_mgml", 9),
    rep("cali_025_mgml", 9),
    rep("cali_050_mgml", 9),
    rep("cali_075_mgml", 9),
    rep("background", 4)
  )
)

raman <- raman_all$subset_analyses(analyses = c(1:36, 46:90, 95:151, 155:160))
raman$add_blank_names(c(rep("blank", 81), rep("blank_e", 9), rep("blank", 9), rep("blank_e", 36), rep("blank", 9)))
```

```{r plot-raman-raw, fig.cap="Raw averaged Raman spectra for each analysis replicate."}
raman$plot_spectra(colorBy = "replicates")
```

```{r raman-workflow}
raman_workflow_settings <- list(
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(),
  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 350))),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2500))),
  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30),
  Settings_normalize_spectra_minmax(),
  Settings_normalize_spectra_blockweight()
)

raman$add_settings(raman_workflow_settings)

raman$print_workflow()
```

```{r raman-run-workflow}
raman$run_workflow()
```

```{r plot-raman-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman$plot_spectra()
```

```{r raman-spectra}
raman$get_spectra()[1:10, ]
```

# Fusion

```{r fusion}
ms_spec_mat <- ms$get_spectra_matrix()
raman_spec_mat <- raman$get_spectra_matrix()
fused_mat <- cbind(ms_spec_mat, raman_spec_mat)
# The matrix is mean centered using the mdatools R package
fused_mat <- mdatools::prep.autoscale(fused_mat, center = TRUE, scale = FALSE)
```

```{r plot-fused, include=FALSE, fig.cap="Fused spectra."}
cl <- StreamFind:::.get_colors(rownames(fused_mat))
fig <- plot_ly()
xVal <- seq_len(length(fused_mat[1, ]))
for (i in seq_len(nrow(fused_mat))) {
  fig <- fig %>% add_trace(
    x = xVal,
    y = fused_mat[i, ],
    type = "scatter", mode = "lines",
    line = list(width = 0.5, color = unname(cl[i])),
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = TRUE
  )
}
xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Bins",
  titlefont = list(size = 12, color = "black")
)
yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Intensity",
  titlefont = list(size = 12, color = "black")
)
fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
fig
```

# Statistic analysis

## Unsupervised

Principle Components Analysis (PCA) can be used for unsupervised evaluation, meaning that we assume that the main (or train) set of analyses are of good quality and we test new analyses to evaluate the proximity to the main set. One could say that it is a supervised approach but in the sense that we assume the quality of the main set. Yet, we are not classifying the possible deviation of the new analyses which would be a clear supervised approach, as we show in the next sub chapter.

```{r sets-show}
main_set <- fused_mat[c(7:9, 16:21, 25:33), ]
rownames(main_set)

test_set <- fused_mat[c(1:6, 10:15, 22:24, 34:42), ]
rownames(test_set)
```

```{r pca-engine, fig.cap="PCA model (native plot from mdatools R package).", fig.height=9}
stat <- StatisticEngine$new(data = main_set)
# the alpha and gamma parameters are used to evaluate the model limits
stat$make_model(Settings_make_model_pca_mdatools(ncomp = 3, alpha = 1E-4, gamma = 1E-5))
# original plot from mdatools R package
plot(stat$model)
```

```{r pca-predict, fig.cap="PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
stat$predict(test_set)
stat$plot_predicted_distances()
```

## Supervised (classification)

The k-nearest neighbors (KNN) can be used for classification of analyses based on pre-classified training analyses.

```{r sets-show-knn}
train_set <- fused_mat[unique(c(grep("1", rownames(fused_mat)), grep("2", rownames(fused_mat)))), ]
train_set <- train_set[order(rownames(train_set)), ]
rownames(train_set)

test_set <- fused_mat[unique(grep("3|4|5|6|7|8|9", rownames(fused_mat))), ]
rownames(test_set)
```

```{r knn-classes}
labels <- c(rep("ass", 2), rep("caff", 2), rep("fiz", 2), rep("fiz_acid", 2), rep("fiz_ass", 2),
  rep("fiz", 2), rep("fiz", 2), rep("fiz_caff", 2), rep("fiz", 7), rep("par", 2))
```

```{r knn-classes-show, echo=FALSE}
kable(data.frame(replicate = rownames(train_set), class = labels), caption = "Class assignment for training set.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r knn-engine}
stat2 <- StatisticEngine$new(data = train_set)
stat2$add_classes(labels)
stat2$prepare_classification(Settings_prepare_classification_knn(k = 3))
stat2$classify(test_set)
stat2$classification_results
```

# Quantification with Raman

```{r ov-raman-conc, include=FALSE}
raman_quant <- raman_all$subset_analyses(analyses = c(19:27, 161:205, 155:157))
raman_quant$add_blank_names(rep("blank", 57))

ov_quant <- copy(raman_quant$get_overview()[, 1:3])
setorder(ov_quant, replicate)
ov_quant$concentration <- as.numeric(c(rep(0, 9), sub(".*?(\\d+).*", "\\1", ov_quant$replicate[10:nrow(ov2)]))) / 10
ov_quant$concentration[ov_quant$replicate %in% "par_3"] <- NA_real_

kable(ov_quant, caption = "Overview of Raman analyses for quantification with respective concentration.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r plot-raman-conc-raw, fig.cap="Raw averaged Raman spectra for each analysis replicate."}
raman_quant$plot_spectra(colorBy = "replicates")
```

```{r raman-conc-workflow}
raman_workflow_settings <- list(
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(),
  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 350))),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2500))),
  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30)
)

raman_quant$add_settings(raman_workflow_settings)

raman_quant$print_workflow()
```

```{r raman-conc-run-workflow}
raman_quant$run_workflow()
```

```{r plot-raman-conc-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman_quant$plot_spectra()
```

## Applying MCR-ALS

The quantification of paracetamol from the Raman spectra is performed using the Multivariate Curve Resolution (MCR) with Alternating Least Squares (ALS) method via integration of the [mdatools](https://mdatools.com/docs/mcr--als.html) R package.

```{r mcrals-show}
stat3 <- StatisticEngine$new(data = raman_quant$get_spectra_matrix())
stat3$add_concentrations(c(1, 2.5, 5.0, 7.5, 15.0, NA_real_)) # add concentrations,with NA for unknown
stat3$make_model(Settings_make_model_mcrals_mdatools(ncomp = 3))
summary(stat3$model)
```

```{r mcrals-resolved-spec, fig.cap="Resolved spectra for each component and the spectra for each analysis replicate (i.e., each concentration)."}
stat3$plot_model_resolved_spectra()
```

```{r mcrals-cont, fig.cap="Contributions for each analysis replicate."}
stat3$plot_model_contributions()
```

The `quantify` method can be used to estimate the concentration of the paracetamol analyses with unknown concentration. The quantification is applied to each component, returning a list with the r square of the linear model and the estimated concentration as `data.table`. Note that the explained variance for each component is important to evaluate the calculated concentration. The best estimation is from the component with the combined explained variance and r square closer to 100 and 1, respectively.

```{r}
stat3$quantify()
```