### 2024-08-01 README.Rmd
Updated settings, articles and core

---
output: github_document
bibliography: inst/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# StreamFind (R package)

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

<p align="center" width="100%">
  <img width="60%" src="man/figures/logo_StreamFind.png" alt="Logo" />
</p>

The StreamFind is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html), which is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de). The development is carried out by the [Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners. The goal of the StreamFind is to develop and assemble data processing workflows for different types of analytical data (e.g., mass spectrometry and spectroscopy) and to apply the workflows in different fields (e.g., environmental and quality studies of the water cycle). StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, to promote standardization of data structure and processing, and to facilitate retrospective data evaluation. The StreamFind platform is aimed at scientists, but also at technicians, due to its comprehensive documentation, its well categorized set of integrated modular functions and its embedded graphical user interface.  
<br>
The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions or collaboration.

```{r, echo=FALSE, out.width = '100%', fig.align='center'}
knitr::include_graphics("man/figures/logo_group.png")
```

## Installation

Pre-requisites for the StreamFind are the [R](https://cran.r-project.org/) software and the [RTools](https://cran.r-project.org/bin/windows/Rtools/) (only applicable for Windows users). RTools is needed for compiling the C++ code used in the StreamFind R package. Assuming that R and RTools are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).

```{r install-streamfind, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("odea-project/StreamFind", dependencies = TRUE)
```

## Other dependencies

The StreamFind depends on other open source software to process different analytical data. For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies. Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites). Consult the documentation for dependencies of other data types.

## Suplementary data

The supplementary [StreamFindData](https://github.com/odea-project/StreamFindData) R package holds the data used in examples and other documentation assets of the StreamFind and can also be installed from the GitHub repository.

```{r install-streamfinddata, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("odea-project/StreamFindData")
```

## Documentation

The documentation and usage examples of the StreamFind R package can be found in the [reference page](https://odea-project.github.io/StreamFind/reference/index.html) and [articles](https://odea-project.github.io/StreamFind/articles/index.html) of the [webpage](https://odea-project.github.io/StreamFind/index.html), respectively.

## References

The StreamFind is open source due to public funding and the extensive contribution from scientific literature as well as existing open source software. Below, we reference the research and software that is used within StreamFind. Please note that each open source software or research that StreamFind uses relies on other contributions. Therefore, we recommend to search within each citation for other contributions.  
<br>

---
nocite: '@*'
---