### 2024-09-17 dev/dev_template_report.qmd
Added annotation of adducts in NTS

---
title: "Short Report on Mass Spectrometry"
author: "Ricardo Cunha"
date: "`r format(Sys.time(), '%d %B, %Y')`"
toc: true
format:
  html:
    grid:
      sidebar-width: 300px
      body-width: 1200px
      margin-width: 300px
      gutter-width: 1.5rem
    other-links:
      - text: StreamFind
        href: https://odea-project.github.io/StreamFind/
        icon: github
    html-math-method: katex
    theme: journal
    code-tools: true
    anchor-sections: true
    code-fold: true
    code-overflow: scroll
    code-line-numbers: false
    code-copy: true
    cap-location: top
    self-contained: true
    embed-resources: true
    toc: true
    toc-location: left
execute:
  warning: false
---

```{r}
#| include: false
options(DT.options = list(
  dom = "lfrtip",
  autoWidth = TRUE,
  pageLength = 10,
  lengthMenu = c(10, 20, 50, 75, 100),
  initComplete = htmlwidgets::JS(
    "function(settings, json) {",
    "$(this.api().table().body()).css({'font-size': '12px'});",
    "$(this.api().table().body()).css({'white-space': 'nowrap'});",
    "}"
  )
))
```

```{r}
#| include: false
library(knitr)
library(kableExtra)
library(data.table)
library(DT)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)
```

```{r}
#| include: false
ms_files <- StreamFindData::get_ms_file_paths()[10:27]
db <- StreamFindData::get_ms_tof_spiked_chemicals_with_ms2()
cols <- c("name", "formula", "mass", "rt", "fragments", "tag")
db <- db[, cols, with = FALSE]
dbis <- db[grepl("IS", db$tag), ]
dbsus <- db[!grepl("IS", db$tag), ]
path <- "C:/Users/apoli/Documents/example_ms_files"
```

# Analyses

```{r}
#| code-fold: show
ms <- MassSpecEngine$new(analyses = ms_files)
```

```{r}
DT::datatable(ms$analyses$info)
```

<br>  
<br>

```{r}
ms$plot_spectra_tic(levels = 1, interactive = TRUE, colorBy = "replicates", title = "Total Ion Chromatograms (TICs)")
```

<br>  
<br>