### 2024-10-27 dev/dev_template_dashboard.qmd

---
title: "Short Dashboard for Mass Spectrometry"
logo: "`r file.path(system.file(package = 'StreamFind', dir = 'extdata'), 'logo_StreamFind.png')`"
format:
  dashboard:
    theme: lux
    self-contained: true
    embed-resources: true
    nav-buttons:
      - icon: github
        href: https://github.com/odea-project/StreamFind
editor_options:
  chunk_output_type: console
---

```{r}
#| include: false

options(DT.options = list(
  bPaginate = TRUE, 
  dom = "lfrtip",
  pageLength = 20,
  lengthMenu = c(10, 15, 20, 25, 50, 75, 100),
  language = list(info = 'Showing page _PAGE_ of _PAGES_')
))
```

```{r}
#| include: false

library(knitr)
library(kableExtra)
library(data.table)
library(DT)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)
```

```{r}
#| include: false

ms_files <- StreamFindData::get_ms_file_paths()[10:27]
db <- StreamFindData::get_ms_tof_spiked_chemicals_with_ms2()
cols <- c("name", "formula", "mass", "rt", "fragments", "tag")
db <- db[, cols, with = FALSE]
dbis <- db[grepl("IS", db$tag), ]
dbsus <- db[!grepl("IS", db$tag), ]
path <- "C:/Users/apoli/Documents/example_ms_files"
```

# Analyses

## Row {.tabset}

::: {.card title="Analyses Overview"}

```{r}
#| code-fold: show

ms <- MassSpecEngine$new(analyses = ms_files)
```

```{r}
DT::datatable(ms$analyses$info)
```

:::

::: {.card title="Total Ion Chromatograms"}

```{r}
ms$plot_spectra_tic(levels = 1, interactive = TRUE, colorBy = "replicates", title = "Total Ion Chromatograms (TICs)")
```

:::

# Processing {scrolling="true"}

## Row {.tabset}

::: {.card title="Find Features"}

Extracted ion chromatograms of Metoprolol-D7 (*m/z* 274.2274) and Carbamazepine-D10 (*m/z* 246.1577) in analysis 124 are shown below.

```{r}
plot_spectra_xic(ms$analyses, analyses = 5, mass = dbis[c(4, 8), ], ppm = 50, sec = 30)
```

```{r}
ffs <- MassSpecSettings_FindFeatures_openms()
ffs
```

```{r}
ms$run(ffs)
```

```{r}
plot_features_count(ms$analyses)
```

:::