### 2025-02-07 dev/demo_paracetamol_fusion.Rmd

---
title: 'Detect Pharmaceutical Quality Deviations with MS, Raman and Statistics'
subtitle: |
  | Fusion of Mass Spectrometry and Raman Spectroscopy Data
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{'Detect Pharmaceutical Quality Deviations with MS, Raman and Statistics'}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  results = "markup",
  comment = ""
)

options(
  DT.options = list(
    dom = "lfrtip",
    autoWidth = TRUE,
    pageLength = 10,
    lengthMenu = c(10, 20, 50, 75, 100),
    initComplete = htmlwidgets::JS(
      "function(settings, json) {",
      "$(this.api().table().body()).css({'font-size': '12px'});",
      "$(this.api().table().body()).css({'white-space': 'nowrap'});",
      "}"
    )
  )
)

main_dir <- "C:/Users/apoli/Documents/example_files/isoft/Paracetamol_subset" # Ricardo laptop
ms_dir <- paste0(main_dir, "/ms")
ms_files <- list.files(ms_dir, pattern = ".mzML$", full.names = TRUE)
raman_dir <- paste0(main_dir, "/raman")
raman_files <- list.files(raman_dir, pattern = ".asc$", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

This article demonstrates how StreamFind, using mass spectrometry (MS) combined with Raman spectroscopy data, can be used to assess quality variation in the pharmaceutical Fizamol, which contains paracetamol as the active ingredient (API). Each product sample was measured by both MS and Raman spectroscopy. The data are pre-processed and fused to be used in unsupervised and supervised statistical analysis for quality assessment.

# Pre-processing

Data is pre-processed using StreamFind's built-in data-specific engines for MS and Raman data. Pre-processing workflows for each data type are performed prior to data fusion to improve statistical analysis.

## MS

```{r ms-setup, include=FALSE}
ms <- MassSpecEngine$new(analyses = ms_files)

ms$analyses$replicates <- c(
  rep("blank_01", 4),
  rep("fiz_01", 3),
  rep("fiz_02", 3),
  rep("fiz_03", 3),
  rep("fiz_04", 3),
  rep("fiz_05", 3),
  rep("fiz_06", 3),
  rep("blank_02", 8),
  rep("fiz_07", 3),
  rep("fiz_08", 3),
  rep("fiz_09", 3),
  rep("fiz_10", 3),
  rep("fiz_11", 3),
  rep("fiz_12", 3),
  rep("fiz_13", 3),
  rep("fiz_14", 3),
  rep("fiz_15", 3),
  rep("fiz_acid_1", 3),
  rep("fiz_acid_2", 3),
  rep("fiz_acid_3", 3),
  rep("fiz_ass_1", 3),
  rep("fiz_ass_2", 3),
  rep("fiz_ass_3", 3),
  rep("fiz_caff_1", 3),
  rep("fiz_caff_2", 3),
  rep("fiz_caff_3", 3),
  rep("fiz_old_1", 3),
  rep("fiz_old_2", 3),
  rep("fiz_old_3", 3)
)

ms$analyses$blanks <- c(
  rep("blank_01", 4 + 9 + 9),
  rep("blank_02", 8 + 9 + 9 + 9),
  rep("blank_01", 9 + 9 + 9 + 9)
)
```

```{r ms-overview, echo=FALSE}
ov <- ms$analyses$info[, 1:3]

ov$description <- c(
  rep("Reference background 1", 4),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Reference background 2", 8),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with acetic acid", 9),
  rep("Paracetamol with aspirin", 9),
  rep("Paracetamol with caffeine", 9),
  rep("Paracetamol stored", 9)
)

kable(ov, caption = "Overview of MS analyses and respective analysis replicate groups and blanks.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

<br>

```{r ms-workflow}
ms_workflow <- list(
  MassSpecSettings_LoadSpectra_StreamFind(levels = 1, mzmax = 300, rtmax = 120),
  MassSpecSettings_BinSpectra_StreamFind(binNames = c("rt", "mz"), binValues = c(5, 2), refBinAnalysis = 1),
  MassSpecSettings_NormalizeSpectra_minmax(),
  MassSpecSettings_AverageSpectra_StreamFind(),
  MassSpecSettings_SubtractBlankSpectra_StreamFind(negativeToZero = TRUE),
  MassSpecSettings_NormalizeSpectra_blockweight()
)

ms$workflow <- ms_workflow

ms$print_workflow()
```

```{r ms-run-workflow}
ms$run_workflow()
```

```{r plot-ms-spectra, fig.cap="Pre-processed spectra for each analysis. Retention time dimension in he created bins is collapsed for each mass-to-charge ratio"}
ms$plot_spectra()
```

## Raman

```{r raman-setup, include=FALSE}
raman <- RamanEngine$new(analyses = raman_files)

raman$analyses$replicates <- c(
  rep("blank_01", 9),
  rep("fiz_01", 3),
  rep("fiz_02", 3),
  rep("fiz_03", 3),
  rep("fiz_04", 3),
  rep("fiz_05", 3),
  rep("fiz_06", 3),
  rep("blank_02", 9),
  rep("fiz_07", 3),
  rep("fiz_08", 3),
  rep("fiz_09", 3),
  rep("fiz_10", 3),
  rep("fiz_11", 3),
  rep("fiz_12", 3),
  rep("fiz_13", 3),
  rep("fiz_14", 3),
  rep("fiz_15", 3),
  rep("fiz_acid_1", 3),
  rep("fiz_acid_2", 3),
  rep("fiz_acid_3", 3),
  rep("fiz_ass_1", 3),
  rep("fiz_ass_2", 3),
  rep("fiz_ass_3", 3),
  rep("fiz_caff_1", 3),
  rep("fiz_caff_2", 3),
  rep("fiz_caff_3", 3),
  rep("fiz_old_1", 3),
  rep("fiz_old_2", 3),
  rep("fiz_old_3", 3)
)

raman$analyses$blanks <- c(
  rep("blank_01", 9 + 9 + 9),
  rep("blank_02", 9 + 9 + 9 + 9),
  rep("blank_01", 9 + 9 + 9 + 9)
)
```

```{r raman-overview, echo=FALSE}
raman_ov <- raman$get_overview()[, 1:3]

raman_ov$description <- c(
  rep("Reference background 1", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Reference background 2", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with acetic acid", 9),
  rep("Paracetamol with aspirin", 9),
  rep("Paracetamol with caffeine", 9),
  rep("Paracetamol stored", 9)
)

kable(raman_ov, caption = "Overview of MS analyses and respective analysis replicate groups and blanks.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r raman-workflow}
raman_workflow <- list(
  RamanSettings_AverageSpectra_native(),
  RamanSettings_SubtractBlankSpectra_StreamFind(),
  RamanSettings_SmoothSpectra_savgol(fl = 21, forder = 4, dorder = 0),
  RamanSettings_DeleteSpectraSection_native(min = -100, max = 350),
  RamanSettings_DeleteSpectraSection_native(min = 1800, max = 2500),
  RamanSettings_CorrectSpectraBaseline_airpls(lambda = 45, differences = 1, itermax = 30),
  RamanSettings_NormalizeSpectra_minmax(),
  RamanSettings_NormalizeSpectra_blockweight()
)

raman$workflow <- raman_workflow

raman$print_workflow()
```

```{r raman-run-workflow}
raman$run_workflow()
```

```{r plot-raman-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman$plot_spectra()
```

# Fusion

Statistical analysis is performed on the merged data for combination of orthogonal information in MS and Raman data.

```{r fusion}
ms_spec_mat <- ms$get_spectra_matrix()
raman_spec_mat <- raman$get_spectra_matrix()
fused_mat <- cbind(ms_spec_mat, raman_spec_mat)
# The fused matrix is mean centered using the mdatools R package
fused_mat <- mdatools::prep.autoscale(fused_mat, center = TRUE, scale = FALSE)
str(fused_mat)
```

# Statistic analysis

## Unsupervised

Principle Components Analysis (PCA) can be used for unsupervised evaluation, that is, we assume that the main (or training) set of analyses is of good quality and we test new analyses to evaluate their closeness to the main set. You could say it is a supervised approach, but in the sense that we assume the quality of the main set. However, we do not classify the possible deviation of the new analyses, which would be a clearly supervised approach, as we show in the next subchapter.

```{r sets-show}
main_set_idx <- c(1:10)

main_set <- fused_mat[main_set_idx, ]
rownames(main_set)

test_set <- fused_mat[-main_set_idx, ]
rownames(test_set)
```

```{r pca-engine, fig.cap="PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
stat <- StatisticEngine$new(analyses = main_set)
# The alpha and gamma parameters are used to evaluate the model limits
stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
plot(stat$model, showText = TRUE)
```

```{r pca-predict, warning=FALSE, fig.cap="PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
stat$predict(test_set)
plot(stat$model, showText = TRUE)
```

## Supervised (classification)

K-Nearest Neighbors (KNN) can be used to classify analyses based on pre-classified training analyses.

```{r sets-show-knn}
train_idx <- unique(c(grep("1", rownames(fused_mat)), grep("2", rownames(fused_mat))))

train_set <- fused_mat[train_idx, ]
train_set <- train_set[order(rownames(train_set)), ]
rownames(train_set)

test_set <- fused_mat[-train_idx, ]
rownames(test_set)
```

```{r knn-classes}
labels <- c(
  rep("fiz", 8),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

```{r knn-engine}
stat2 <- StatisticEngine$new(analyses = train_set)
stat2$add_classes(labels)
stat2$run(StatisticSettings_MakeModel_knn(k = 3))
stat2$predict(test_set)
stat2$model$prediction$results
```