### 2025-04-19 dev/demo_paracetamol_quantification_mcrals.Rmd

---
title: 'Pharmaceutical API Quantification with Raman and MCR'
subtitle: |
  | Raman Spectroscopy Data and Multivariate Curve Resolution with Alternating Least Squares (MCR-ALS)
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Pharmaceutical API Quantification with Raman and MCR}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  results = "markup",
  comment = ""
)

raman_dir <- "C:/Users/apoli/Documents/example_files/isoft/Paracetamol_raman_quantification"
raman_files <- list.files(raman_dir, pattern = ".asc$", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>
<br>

***

This article demonstrates how StreamFind can be used to quantity the active product ingredient (API) in Fizamol (i.e., 
paracetamol) from Raman data and using the Multivariate Curve Resolution with Alternating Least Squares (MCR-ALS) 
statistical method. A calibration curve is acquired with pure paracetamol diluted in water at different concentrations.
Two Fizamol tablets (*fiz_a* and *fiz_b*) are diluted in water to a concentration of approx. 10 mg/ml and pure 
paracetamol (*par*) at 10 mg/ml are measured for quantification. Water is used as a blank.

```{r raman-setup}
# Creates the engine for Raman data
raman <- RamanEngine$new(analyses = raman_files)

# Modifies the replicate names
raman$analyses$replicates <- c(
  rep("blank", 9),
  rep("fiz_a", 9),
  rep("fiz_b", 9),
  rep("par", 3),
  rep("cali_100_mgml", 9),
  rep("cali_150_mgml", 9),
  rep("cali_010_mgml", 9),
  rep("cali_025_mgml", 9),
  rep("cali_050_mgml", 9),
  rep("cali_075_mgml", 9)
)

# Modifies the blank names
raman$analyses$blanks <- c(rep("blank", length(raman$analyses)))
```

<br>
<br>

```{r raman-overview, echo=FALSE}
ov_quant <- raman$get_overview()[, 1:3]
data.table::setorder(ov_quant, replicate)
ov_quant$concentration <- as.numeric(c(rep(0, 9), sub(".*?(\\d+).*", "\\1", ov_quant$replicate[10:63]), rep(0, nrow(ov_quant) - 63))) / 10
ov_quant$concentration[ov_quant$replicate %in% c("par", "fiz_a", "fiz_b")] <- NA_real_
kable(ov_quant, caption = "Overview of Raman analyses and respective analysis replicate groups and blanks.") %>%
  kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "400px")
```

<br>
<br>

```{r plot-raman-conc-raw, fig.cap="Raw Raman spectra for each analysis replicate."}
raman$plot_spectra(colorBy = "replicates")
```

# Pre-processing Raman Data

The data is pre-processed using a workflow that includes the following steps:

1. Averaging the spectra for each analysis replicate.
2. Subtracting the blank spectra.
3. Smoothing the spectra using the Savitzky-Golay method.
4. Deleting the spectra sections that are not relevant for the analysis.
5. Correcting the spectra baseline using the Asymmetric Least Squares (airPLS) method.

The methods ca be applied one at the time or used to bluid a Workflow object that will apply all methods in sequence 
within engine, as shown below.

```{r raman-conc-workflow}
# Creates the workflow for Raman data
# with an ordered list of processing settings
raman_workflow <- list(
  RamanSettings_AverageSpectra_native(),
  RamanSettings_SubtractBlankSpectra_StreamFind(),
  RamanSettings_SmoothSpectra_savgol(fl = 7, forder = 4, dorder = 0),
  RamanSettings_DeleteSpectraSection_native(min = -100, max = 350),
  RamanSettings_DeleteSpectraSection_native(min = 1800, max = 2500),
  RamanSettings_CorrectSpectraBaseline_airpls(lambda = 11, differences = 1, itermax = 30)
)

# Adds the workflow to the engine
raman$workflow <- raman_workflow

raman$print_workflow()
```

```{r raman-conc-run-workflow}
# Runs the workflow
raman$run_workflow()
```

```{r plot-raman-conc-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman$plot_spectra()
```

## Applying MCR-ALS for Quantification

The quantification of paracetamol from the pre-processed Raman spectra is performed using the Multivariate Curve 
Resolution (MCR) with Alternating Least Squares (ALS) method via integration of the [mdatools](https://mdatools.com/docs/mcr--als.html) R package.

```{r}
# Gets a matrix with spectra from the Raman engine
raman_matrix <- raman$get_spectra_matrix()
str(raman_matrix)
```

```{r mcrals-show}
# Creates the statistic engine for the statistic analysis
stat <- StatisticEngine$new(analyses = raman_matrix)

stat
```

```{r}
# Applies the MCR-ALS method with the run method of the engine
stat$run(StatisticSettings_MakeModel_mcrals_mdatools(ncomp = 2))
```

```{r mcrals-resolved-spec, fig.cap="Overview plot for the MCR-ALS model"}
plot(stat$model)
```

```{r}
# Applies the quantification method with the run method of the engine
# Concentrations with NA for unknown values
stat$run(
  StatisticSettings_Quantify_mcrals(
    concentrations = c(1, 2.5, 5.0, 7.5, 10, 15.0, NA_real_, NA_real_, NA_real_)
  )
)
```

The MCR-ALS was applied with two components and the quantification results are shown below. The component with the 
highest combined explained variance and R<sup>2</sup> values for the calibration curve should be used for quantification. 
The analysis *par* should have about 10 in mg/ml of paracetamol while the analyses *fiz_a* and *fiz_b* should a little 
less than 10 mg/ml due to presence of formulation ingredients and possible loss of API during sample preparation.

```{r}
# Gets the quantification results
stat$quantification@quantities
```